---
title: "Regression Models Course Project"
output: pdf_document
---

## Executive Summary

Per the analysis below, the relationship between mpg and the available data is best catpured using a model that regresses weight and horse power against mpg. For a given horse power and weight, manual transmission have a higher mpg. The incremental effect of a manual transmission is approximatley 2.08 additional miles/gallon.

## Synopsis
This analysis will explore the relationships between 11 variables for 32 automobiles collected by Motor Trend magazine in 1974 and answer the following questions:

* Is an automatic or manual transmission better for MPG?
* What is the MPG difference between automatic and manual transmissions?

## Loading and Reviewing Raw Data

The dataset used is the mtcars dataframe in the R datasets package.

```{r echo=FALSE}
data(mtcars)
```

A description of the variables in this dataset can be found [here](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html).

## Exploration

A pairs comparison of the variables in the dataset can be found [here](https://github.com/EnriqueElGato/RegressionModels/blob/master/appendix1.pdf). Some variable appear highly correlated with mpg.

## Model Selection

### Initial Transmision Model

A model must balances bias and variability in model estimates. Below is summary of a regression using only transmission. It captures a low percentage of the observered variation in mpg values. 

```{r,results='asis',echo=FALSE,warning=FALSE,message=FALSE}
library(xtable,quietly=TRUE,warn.conflicts = FALSE)
fit1 <- lm(mpg ~ factor(am) ,data=mtcars)
s<-summary(fit1)
f<-s$fstatistic
d<-data.frame(Adj.R.Sqred=s$adj.r.squared,F.Stat=f[1],Resid.Std.Err=s$sigma)
cF <-as.character(fit1$call$formula)
cF <- c(cF[2],cF[1],cF[3:length(cF)])
name <- paste(paste(cF[1],cF[2]),cF[3:length(cF)],collapse=" + ")
dimnames(d)[[1]] <- name
xt<-xtable(d,caption="Transmission Only Model Fit",digits = 4)
print(xt,type = "latex",zero.print = ".",comment=FALSE)
```

### Selection Of Variables

To selectively identify model variables, I with the variable with the highest correlations with mpg, weight. 

```{r,results='asis',echo=FALSE,warning=FALSE,message=FALSE} 
df <- data.frame(matrix(rep(mtcars[,1],10),ncol=10))
names(df) <- names(mtcars[,2:11])
tab <- mapply(cor,df,mtcars[,2:11])
mpgCors <-matrix(tab,ncol=10)
dimnames(mpgCors) <- list(c("mpg"),names(mtcars[,2:11]))

wtCors <- with(mtcars,
        matrix(c(cor(wt,cyl),cor(wt,disp),cor(wt,hp)),nrow=1)
)
dimnames(wtCors) <-list(c("wt"),c("cyl","disp","hp"))
xt <- xtable(mpgCors,caption="Mpg Correlations")
print(xt,type = "latex",zero.print = ".",comment=FALSE)
```

I then examined the next three hightest correlations and used  one with the smallest correltain with weight, horse power. 

```{r,results='asis',echo=FALSE,warning=FALSE,message=FALSE} 
xt <- xtable(wtCors,caption="Weight Correlations")
print(xt,type = "latex",zero.print = ".",comment=FALSE)
```

### Analysis of Variance 

To check if any additional variables should be added, an analysis was done to see the impact of adding one additional varible. None of the other variables had statistically significant p-values. Only mpg, weight and horse power will be used in the model to analyize transmission impact.

```{r,echo=FALSE,warning=FALSE,message=FALSE}
fbind<-function(f,tab,doANOVA){
        s<-summary(f)
        fstat<-s$fstatistic
        anova.PV <- NA
        if(doANOVA & !is.null(lastFit)){
                a <- anova(lastFit,f)
                anova.PV <- a$"Pr(>F)"[2]
        } 
        d<-data.frame(Adj.R.Sqred=s$adj.r.squared,F.Stat=fstat[1],Resid.Std.Err=s$sigma,anova.P.Value=c(anova.PV))
        cF <-as.character(f$call$formula)
        cF <- c(cF[2],cF[1],cF[3:length(cF)])
        name <- paste(paste(cF[1],cF[2]),cF[3:length(cF)],collapse=" + ")
        dimnames(d)[[1]] <- name
        if((!is.na(anova.PV) & anova.PV < 0.05) | (is.null(lastFit) & doANOVA) ){
                lastFit <<- f
        }  
        if(is.null(tab)){
                d        
        }else{
                rbind(tab,d)
        }

}
f0 <- lm(mpg ~ .,data=mtcars)
f1 <- lm(mpg ~ wt,data=mtcars)
f2 <- lm(mpg ~ wt + hp,data=mtcars)
f3 <- lm(mpg ~ wt + hp + factor(am),data=mtcars)
f4 <- lm(mpg ~ wt + hp + factor(cyl),data=mtcars)
f5 <- lm(mpg ~ wt + hp + drat,data=mtcars)
f6 <- lm(mpg ~ wt + hp + qsec,data=mtcars)
f7 <- lm(mpg ~ wt + hp + factor(vs),data=mtcars)
f8 <- lm(mpg ~ wt + hp + factor(gear),data=mtcars)
f9 <- lm(mpg ~ wt + hp + factor(carb),data=mtcars)
f10 <- lm(mpg ~ wt + hp + disp,data=mtcars)
lastFit <- NULL
tab<-fbind(f0,NULL,FALSE)
tab<-fbind(f1,tab,TRUE)
tab<-fbind(f2,tab,TRUE)
tab<-fbind(f3,tab,TRUE)
tab<-fbind(f4,tab,TRUE)
tab<-fbind(f5,tab,TRUE)
tab<-fbind(f6,tab,TRUE)
tab<-fbind(f7,tab,TRUE)
tab<-fbind(f8,tab,TRUE)
tab<-fbind(f9,tab,TRUE)
tab<-fbind(f10,tab,TRUE)
```

```{r,results='asis',echo=FALSE,warning=FALSE,message=FALSE}
xt<-xtable(tab,digits = 4)
print(xt,type = "latex",zero.print = ".",comment=FALSE)
```

##Analysis Of Transmission

For the transmission variable, a value of 0 means automatic and 1 means manual. The intercept captures the mpg for automatic transmission. The factor(am)1 coefficient captures manual transmission marginal impact.See the Executive Summary for remaining analysis.


```{r,results='asis',echo=FALSE,warning=FALSE,message=FALSE}
f3 <- lm(mpg ~ wt + hp + factor(am),data=mtcars)
s<-summary(f3)
xt<-xtable(s,digits = 4)
print(xt,type = "latex",zero.print = ".",comment=FALSE)
```

A plot of the residuals for this model can be found [here](https://github.com/EnriqueElGato/RegressionModels/blob/master/appendix2.pdf).
